---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
  word_document: default
---

## Loading and preprocessing the data
```{r, preprocessing}
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
unzip("activity.zip")
activity <- tbl_df(read.csv("activity.csv"))
activity$date <- ymd(activity$date)
# Sum of steps per day
per_day <- activity %>% 
      group_by(date) %>%
      summarise(sum(steps,na.rm=T))
colnames(per_day) <- c("date","steps")
# Mean of steps per interval
per_interval <- activity %>%
      group_by(interval) %>%
      summarise(mean(steps,na.rm=T))
colnames(per_interval) <- c("interval","steps_mean")
```


## What is mean total number of steps taken per day?
```{r}
qplot(steps,data=per_day)
mean = round(mean(per_day$steps,na.rm=T),0)
median = median(per_day$steps,na.rm=T)
```

The mean total number of steps taken per day is `r signif(mean,5)` and the median total number of steps taken per day is `r median`

## What is the average daily activity pattern?
```{r, plotting & maximum}
qplot(per_interval$interval,per_interval$steps_mean,
      xlab="Interval",ylab="Steps",geom="line")
max_interval = subset(per_interval,steps_mean==max(per_interval$steps_mean))
```

The 5-minute interval with the maximum number of steps (`r max_interval[2]`) was minute `r max_interval[1]` to `r max_interval[1] +5`

There are `r sum(is.na(activity$steps))` missing rows in our dataset.

## Imputing missing values
Impute missing values by using the mean for the 5 minute interval.
```{r ,imputation}
activity <- merge(activity,per_interval,by="interval",all.x=T)
activity$steps <- ifelse(is.na(activity$steps),activity$steps_mean,activity$steps)
```

```{r , summary_stats}
per_day <- activity %>% 
      group_by(date) %>%
      summarise(sum(steps,na.rm=T))
colnames(per_day) <- c("date","steps")
per_interval <- activity %>%
      group_by(interval) %>%
      summarise(mean(steps,na.rm=T))
colnames(per_interval) <- c("interval","steps_mean")
qplot(steps,data=per_day)
mean = round(mean(per_day$steps,na.rm=T),0)
median = median(per_day$steps,na.rm=T)
# Can't seem to change the mean and median values from showing up in scientific notations
```

The mean total number of steps taken per day after imputation is `r round(mean,0)` and the median total number of steps taken per day is `r round(median,0)`


## Are there differences in activity patterns between weekdays and weekends?
```{r , weekday/weekend}
activity$day = weekdays(activity$date)
## If the "day" begins with an "S" character, code as weekend, else weekend.
activity$weekday = as.factor(ifelse(grepl("^S",activity$day)==T,"Weekend","Weekday"))
```

```{r,cache=FALSE}
summary_activity = activity %>%
      group_by(interval,weekday) %>%
      summarize(steps = mean(steps))
ggplot(summary_activity,aes(interval,steps)) + geom_line() + facet_grid(. ~ weekday)
```


